# Credit

## GridCAT

This example pipeline uses the data that accompanies the GridCAT toolbox.

Here is the reference to the toolbox:  

>Stangl, M., Shine, J., & Wolbers, T. (2017). The GridCAT: A Toolbox for Automated Analysis of Human Grid Cell Codes in fMRI. Frontiers in Neuroinformatics, 11. https://doi.org/10.3389/fninf.2017.00047

The data can be downloaded from [here](https://www.nitrc.org/projects/gridcat/).

## Publishing on GitHub Pages

To be able to host the HTML output using GitPages, we need to include a file called ```.nojekyll``` in the docs folder where the HTML documentation is stored.

Let's create this file if it doesn't exist yet.

```{r}

if (!file.exists(here("docs", ".nojekyll"))){
  file.create(here("docs", '.nojekyll'))
}  
```

## List of R packages
Here is a list of packages used in the analysis and references to them. It is created using the [grateful package](https://github.com/Pakillo/grateful).  

Unfortunately, not all packages provide all information, so there are a few warning messages. The code below generates a markdown file with references that we read in and print below. 

```{r create_bibliography}

if(!file.exists(here("code", "apa-6th-edition.csl"))){
  grateful::get_csl("apa-6th-edition")
  file.copy(from=here("apa-6th-edition.csl"), to=here("code", "apa-6th-edition.csl"))
  
}

pkgs <- grateful::cite_packages(citation.style = "apa-6th-edition", output = "table")
knitr::kable(pkgs)
```


## Session Info

Lastly, we run session info, using the version from the devtools package.

```{r session_info}
sessionInfo()
```


## Bibliography
